### 2019-11-06 latex/vhcs.tex
Added a VHCs tex file
Started writing up content about virtual holonomic constraints. We will
need to move the lagrangian stuff from vnhcs to here, as well as the
conversion from lagrangian to hamiltonian stuff.

%! TEX root = main.tex

%/========== Virtual Holonomic Constraints ==========/%
\section{Virtual Holonomic Constraints}\label{sec:vhcs}

\subsection{Lagrangian Framework Review}%
\label{sub:lagrangian_vhcs}

\begin{defn}\label{defn:vhc_order_k}
   A \textbf{virtual holonomic constraint (VHC) of order k} is a smooth relation 
   \( h : \mathcal{Q} \rightarrow \mathbb{R}^k\) where
   \( \text{rank}(dh_q) = k \, \forall q \in h^{-1}(0)\) and where there exists
   a smooth feedback controller \(\tau(q,\dot{q})\) such that the constraint
   manifold
   \[
      \Gamma = \left\{ (q,\dot{q}) \in T_q\mathcal{Q} \vert h(q) = 0, \, 
      dh_q \dot{q} = 0\right\}
   \]
   is positively invariant.
\end{defn}

\begin{defn}\label{defn:vhc_regular}
   A VHC is \textbf{regular} if system \textbf{TODO:
   reference to E-L mechanical system} with output \(e = h(q)\) has vector
   relative degree \(\{2,2,\ldots,2\}\).
   Equivalently, a VHC is regular if 
   \[
      dh_q D^{-1}(q)B(q)
   \]
   is nonsingular everywhere on \(\Gamma\).
\end{defn}

%/========== /Virtual Holonomic Constraints ==========/%
% vim: set tw=80 ts=3 sw=3 sts=0 et ffs=unix :