### 2020-04-09 latex/ogs/main.tex
Rough Draft update: changed the goal to transitioning between VNHCs

%! TEX program = lualatex

\documentclass[a4paper,12pt]{article}

%/========== Preamble ==========/%
%----------- Set Main Font ----------%
\usepackage{times} % Times New Roman

%----------- Choose Margins ----------%
\usepackage[a4paper,margin=1.87cm]{geometry}

%---------- Bibliography Style ----------%
\usepackage[style=ieee,backend=biber]{biblatex}
\addbibresource{bib.bib}

%----------- Math Packages -----------%
\usepackage{mathtools}

%---------- Remove indenting from paragraphs ----------%
% Also makes each new paragraph start with an vertical space.
\usepackage[parfill]{parskip}

%/========== Main Document ==========/%
\begin{document}

%---------- Title -----------%
\title{Ontario Graduate Scholarship: Plan of Study}
\author{Adan Moran-MacDonald}
\date{}
\maketitle

%---------- Main Content ----------%
\begin{large}
\textbf{Background}
\end{large}

When a gymnast is performing a swing-up routine on a bar, they do not kick their
feet at specified time intervals. Rather, a more natural human motion is to
swing their legs based on their current speed and position. In doing so, they
are able to generate enough momentum to perform backflips on the bar. When
controlling mechanical systems, standard practice is to have a robot's actuators
track a position dependent on time. Much like the gymnast, this tends to
produce unnatural motion. Furthermore, it is liable to timing errors if the
motion is interrupted in any way. The more recent techniques of virtual constraints 
can rectify these issues and produce realistic, biologically plausible motion.

Mechanical systems can be constrained in two ways. Holonomic constraints
restrict position (for instance, car tires remain a fixed distance apart), while
nonholonomic constraints restrict velocities (such as how those tires cannot
slide sideways). Previous research has shown that one can enforce holonomic
constraints through a robot's actuators \cite{vhcs_for_el_systems}. These are
called ``virtual holonomic constraints" (VHCs). VHCs have been used to control
walking robots \cite{stable-walking}, autonomous bicycles \cite{bicycle},
gymnastics robots \cite{xingbo-thesis}, and snake robots \cite{snake-robot}
among others. These VHCs are a special case of the more general virtual
nonholonomic constraints (VNHCs); in my master's thesis \textbf{TODO: HOW TO
CITE?} I developed a general theory for enforcing VNHCs and designed
one which injects energy into the gymnastics robot created by Wang
\cite{xingbo-thesis}, thereby enabling it to perform backflips on a bar.

\begin{large} \textbf{Proposal} \end{large}

Mohammadi \textit{et.al.} showed that a single VHC can result in several motion
patterns (also called ``orbits"), from which one specific motion can be chosen
if the robot is initialized with a configuration that is ``close enough” to it
\cite{manfredi-orbit-stabilization}. However, their method does not work for all
starting configurations of the robot. We propose to extend this work and 
investigate a method to stabilize and transition between two VNHCs, which should
also allow one to stabilize an orbit from any starting configuration.

VNHCs have been studied for application to human-robot cooperation
\cite{vnhc-human-robot-coop} and walking robots \cite{vnhc-biped-robot}. Horn
\textit{et.al.} have also examined properties of VNHCs for general mechanical
systems \cite{hybrid_zero_dynamics_bipedal_nhvcs} and applied their results
to stabilizing a chosen gait on walking robots. 
Our approach is novel since we will develop techniques to transition between  
VNHCs on general mechanical systems while maintaining safety constraints, 
rather than stabilizing a single VNHC.

We will first study mathematical conditions under which on can transition
between two VNHCs. These conditions will allow us to create a control algorithm
to switch between motion patterns. After completing this mathematical
formulation, we will test the theory on the gymanstics robot built by Wang
\cite{xingbo-thesis} by designing a gymnastics routine which is performed 
by transitioning between different VNHCs.

The advancement in mathematics from this research will enable improvements to
control of other robotic systems. The ability to transition between complex
constraints will allow for more natural expressive motion in many autonomous
systems and may become a standard technique for controlling biologically
inspired robots.

%---------- Bibliography ----------%
\printbibliography
\end{document}

% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :