### 2020-06-04 latex/abstract.tex
Create a basic skeleton
This commit creates the basic skeleton for the preliminary pages -
abstract, dedication, acknowledgements, title, toc, as well as the
bibliography.

Todo is still all the chapters

%! TEX root = main.tex

%% (At most 150 words for M.Sc. or 350 words for Ph.D.)
\begin{abstract}
	\textbf{TODO}: Fill in the abstract
\end{abstract}
% vim: set ts=3 sw=3 sts=0 et ffs=unix :