### 2020-06-08 latex/conclusion.tex
Finalize the first skeleton of the thesis
This commit finishes the skeleton of the thesis, with all sections and
headings as desired. There is currently no content here.

%! TEX root = main.tex

%/========== Conclusion ==========/%

\chapter{Conclusion}\label{ch:conclusion}
% TODO: Summarize the work

\section{Limitations of this Work}

\section{Future Research}

%/========== /Conclusion ==========/%
% vim: set ts=3 sw=3 sts=0 et ffs=unix :