### 2020-07-30 latex/symbols.tex
Major edits to VNHC
Added several new VNHC-related references, updated the whole chapter of
VNHCs, and implemented my new notation using the matrix kronecker
product. Also added further justfication as to why we have our own
definition of VNHCs, which needs more work. Next, we need to update the
proof on why a VNHC is regular iff it is h(q,pu) and satisfies that
matrix rank definition.

%! TEX root = main.tex

%/========== Symbols ==========/%

% Set the heading so it appears in the Table of Contents
\addcontentsline{toc}{chapter}{List of Symbols}
\chapter*{List of Symbols}

% Define a table for the symbols using the booktabs convention
\noindent
\begin{table}[h!]
\resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{ll}
   \toprule
   Symbol & Definition \\
   \midrule
   \(\n\) & The index set \(\{1,\ldots,n\}\) of natural numbers up to \(n\). \\
   \(\R^n\) & Real numbers in \(n\) dimensions. \\
   \(\Rt{T}\) & Real numbers modulo \(T > 0\), with \(\Rt{\infty} = \R\). \\
   \(\mathbb{S}^1\) & The unit circle, equivalent to \(\Rt{2\pi}\). \\
   \(\mathcal{Q}\) & The configuration manifold of a system. \\
   \(\R^{n \times m}\) & The space of real-valued matrices with \(n\) rows and \(m\) columns. \\
   \(\Id{n}\) & The \(n \times n\) identity matrix. \\
   \(\Zmat{n \times m}\) & The \(n \times m\) matrix of all zeros. \\
   \(M_i\) & If \(M\) is a vector, the \(i\)th element of \(M\). \\
         & If \(M\) is a matrix, the \(i\)th column of \(M\). \\
   \(M_{i,j}\) & The value of row \(i\), column \(j\) for the matrix \(M\). \\
   \(\dot{x}\) & Derivative of \(x\) with respect to time \(t\). \\
   \(\nabla_v F\) & If \(F\) is \(\R\)-valued, the gradient of \(F\) with respect to \(v\). \\
            & If \(F : \R^m \rightarrow \R^{n\times n}\), the block matrix gradient \((\pdiff{F}{v_1},\ldots,\pdiff{F}{v_m}) \in \R^{nm \times n}\). \\
   \(dF_v\) & Total differential (Jacobian) of \(F\), equivalent to \((\nabla_v F)\tpose\). \\
   \(\Hess F\) & If \(F : \R^n \rightarrow \R\), the \(n \times n\) Hessian matrix of double derivatives of \(F\). \\
               & If \(F : \R^n \rightarrow \R^k\), the block matrix \((\Hess F_1, \ldots, \Hess F_k) \in \R^{n \times nk}\). \\
   \(\partial_v \partial_w F\) & Derivative matrix of \(F : \R^n \times \R^m \rightarrow \R\), with \((i,j)\) element \(\ppdiff{F}{v_i}{w_j}\). \\
   \(\delta_{i,j}\) & The Kronecker delta: 1 if \(i = j\) and 0 otherwise. \\
   \(\otimes\) & The matrix kronecker product (see Appendix \ref{appendix:matrix-kron-prod}). \\
   \bottomrule
\end{tabular}

}% resizebox
\end{table}

%---------- Nomenclature ----------%
%\chapter*{Nomenclature}
%
%% Define a table for the symbols using the booktabs convention
%\noindent
%\begin{table}[h!]
%\resizebox{\textwidth}{!}{%
%\centering
%\begin{tabular}{ll}
%   \toprule
%   Acronym & Meaning \\
%   \midrule
%   DOF & Degrees of freedom
%   EOC & Equation of constraint \\
%   EOM & Equation of motion \\
%   ODE & Ordinary differential equation \\
%   VNHC & Virtual nonholonomic constraint \\
%   WLOG & Without loss of generality \\
%   \bottomrule
%\end{tabular}
%
%}% resizebox
%\end{table}
%/========== /Symbols ==========/%
% vim: set ts=3 sw=3 sts=0 et ffs=unix :