### 2020-12-19 latex/introduction.tex
First edit of the intro, pre-lit review

%! TEX root = main.tex

%/========== Introduction ==========/%

\chapter{Introduction}
% TODO: Introduce the thesis with a short, sweet, intuitive approach.
% No math, just whet the reader's appetite for what we are doing. What are the motivating examples? Why do we study what we study? Why choose VNHCs over other methods?
Consider a child on a standing swing.
Naturally, they push off the ground and proceed to alternate between standing
and squatting. 
This allows them to swing higher and faster over time, until they have reached
their express goal of going as high and fast as possible.

Now imagine that the child is a robot, and their creator is teaching them to
behave like a real boy.
If the roboticist had studied classical control theory, they would plot the leg
motion of a human child as a trajectory over time and tell the robot to
synchronize with this trajectory.
In an ideal world, this technique would work perfectly; unfortunately, this
control approach is entirely unnatural. 
Humans on swings do not have an internal stopwatch.
Rather, they adjust their leg position based on the current angle of the swing
along with their direction of motion, as in Figure \ref{fig:swing-pos-vel}.
This position-velocity adjustment allows humans to correct for external
disturbances (like strong winds or overly enthusiastic parents).
A person will squat at the peak of the swing and stand at the bottom.
Even if the robot could perfectly track a time-based trajectory, an external
disturbance may swap the order of squatting and standing to the point that
the swing slows down rather than speeding up.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{images/swing_pos_vel.png}
    \caption{A person on a standing swing will squat as they approach the
        bottom of the swing path, and stand after they pass it until they hit
        the peak of their swing. Image taken and modified from
    \cite{pumping_swing_standing_squatting}.}
    \label{fig:swing-pos-vel}
\end{figure}

One recent control technique known as \textit{virtual holonomic constraints}
forces a robot's actuators to track a function of position rather than time.
Sadly, this will not work because one needs velocity to determine when
they have reached the peak of their swing.
In other words, the robot needs to move according to a function of both position
and velocity if it has any hope of behaving like a real person.

These position-velocity controllers are the focus of this thesis: they are called
\textit{virtual nonholonomic constraints}.

\section{Literature Review}

% TODO: write a brief lit review on energy injection and VNHCs.

\section{Statement of Contributions}
Here are the contributions of this thesis.
\begin{itemize}[label={}]
   \item \textbf{Chapter \ref{ch:vnhcs}} The development of the framework of
      virtual nonholonomic constraints.
      This includes the definition of simply actuated mechanical systems, virtual
      nonholonomic constraints, regular constraints, and energy injection.
      Theorem \ref{thm:vnhc-regularity} yields a characterization of regularity, 
      while Theorem \ref{thm:zero-dynamics} provides the constrained dynamics
      for a certain class of systems.
   \item \textbf{Chapter \ref{ch:vlp}} An application of virtual nonholonomic
      constraints to the variable-length pendulum, based on a pumping technique
      used by children on standing swings.
      The chapter culminates in Theorem \ref{thm:vlp-energy-stabilization},
      which guarantees that a certain class of VNHCs will inject energy into
      this system.
   \item \textbf{Chapter \ref{ch:acrobot}} An application of virtual
      nonholonomic constraints to the acrobot, where we develop a constraint
      by examining the intuitive motion performed by human gymnasts.
      The chapter ends with Theorem \ref{thm:acrobot-energy-stabilization},
      which proves that the constraint injects energy into the acrobot.
\end{itemize}

\section{Organization of the Thesis}
The thesis is laid out as such: 
in Chapter \ref{ch:vnhcs} we cover the requisite background on Hamiltonian
mechanics, after which we develop the main theory of virtual nonholonomic
constraints;
in Chapter \ref{ch:vlp} we reformulate a time-optimal energy-injection strategy
for the variable-length pendulum as a virtual nonholonomic constraint, and
prove that the pendulum will gain enough energy to rotate around the bar with
arbitrary speed;
and in Chapter \ref{ch:acrobot} we find a virtual nonholonomic constraint 
which enables the acrobot to kick its legs like a gymnast until it is
performing backflips on a horizontal bar.
Finally, we show experimental results for the acrobot which confirm that the
theory works in the real world.

%/========== /Introduction ==========/%
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :