### 2020-12-19 latex/conclusion.tex
First version of Intro and Conclusion
I commented out the rest of the thesis so that I could focus on editing
these. Make sure to put the rest of them back in!

%! TEX root = main.tex

%/========== Conclusion ==========/%

\chapter{Conclusion}\label{ch:conclusion}

This thesis has shown the utility of VNHCs as a means of injecting energy into
mechanical systems.
In Chapter \ref{ch:vnhcs} we developed the framework of VNHCs for underactuated
mechanical systems, with a focus on VNHCs for simply actuated Hamiltonian
systems.
We then applied this framework to two benchmark systems: the variable-length
pendulum and the acrobot.
In Chapter \ref{ch:vlp} we proved that a certain class of VNHCs will always
inject energy into the VLP, while in Chapter \ref{ch:acrobot} we designed a
constraint inspired by gymnastics and proved it injects energy into the
acrobot.
In the end, we demonstrated that virtual nonholonomic constraints are capable of
injecting and dissipating energy in a robust manner, all while producing
realistic biological motion.

\section{Limitations and Future Research}
Our VNHC framework is relies on the following assumptions:
\begin{enumerate}
    \item The input matrix \(B(q) \equiv B \in \R^{n \times k}\) is constant and
        full rank.
    \item The input matrix has a left-annihilator 
        \(B^\perp \in \R^{(n-k)\times n}\). 
    \item The annihilator matrix \(B^\perp\) is right semi-orthogonal.
    \item The inertia matrix of the system satisfies 
        \(\nabla_{q_u}M(q) = \Zmat{n(n-k) \times n}\).
    \item On the constraint manifold \(\Gamma\), one can solve for \(q_a\) as a
        function of \((q_u,p_u)\).
\end{enumerate}
If any of these assumptions are not satisfied, one may not be able to find the
constrained dynamics in simply actuated coordinates, and the results may not
hold.
In particular, the theoretical guarantees of this thesis do not hold for systems
with physical nonholonomic constraints like friction, nor does it apply to vehicles
with wheels.

This guides us to the following research directions. 
First, one might consider relaxing the assumptions on the input and inertia
matrices to widen the class of systems to which VNHCs can be applied.
Second, one might forgo the assumption that \(q_a\) is solvable on the
constraint manifold; instead, one could borrow from the VHC literature
and parameterize the constraint manifold with a different set of
coordinates than \((q_u,p_u)\).
Third, one might consider abstract mechanical systems with dynamics
\[
    \begin{cases}
        \dot{q} = \Minv(q) p
        , \\
        \dot{p} = -\pdmat - \nabla_q V(q) + Q_{nh}(q,p) + B(q)\tau
        ,
    \end{cases}
\]
where \(Q_{nh}(q,p) \in \R^n\) are the equations for physical nonholonomic
constraints like friction.
Fourth, one can investigate a smoother mechanism for energy stabilization, where
a VNHC is explicitly designed to inject energy below some energy level, and
dissipate energy above it.
For example, this might involve a mechanism for smoothly transferring between 
VNHCs while maintaining certain safety requirements.

Finally, the proof of Theorem \ref{thm:acrobot-energy-stabilization} 
in Chapter \ref{sec:acrobot-proof} opens a door to the idea of VNHC generation. 
One might search for suitable conditions on a VNHC whereby the Poincar\'{e}
sections for oscillations and rotations are increasing.
Then, one can extend Otsason's ``virtual constraint
generator" \cite{vcg} with the goal of generating regular VNHCs which satisfy
these energy injection/dissipation conditions.
In this way, one could automatically generate VNHCs which stabilize a
desired energy level.

%/========== /Conclusion ==========/%
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :