### 2020-12-21 latex/abstract.tex

%! TEX root = main.tex

%/========== Abstract ==========/%
%% (At most 150 words for M.Sc. or 350 words for Ph.D.)
\begin{abstract}
   This thesis develops a theoretical foundation for the framework of virtual
   nonholonomic constraints, which are relations between the generalized
   coordinates and generalized momenta of a mechanical system that can be
   enforced via feedback control.
   The theory is applied towards energy injection for two standard underactuated
   systems: the variable-length pendulum and the acrobot. 
   Virtual nonholonomic constraints are designed for each system by examining
   human motion, and energy injection properties of these constraints are proven
   rigorously.
   The acrobot constraint is tested on a real-world acrobot, demonstrating
   highly effective energy-injection properties and robustness to a variety of
   external disturbances.
\end{abstract}

%/========== /Abstract ==========/%
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :