### 2020-12-22 latex/acknowledgements.tex
First version of Manfredi's acknowledgement, not done yet

%! TEX root = main.tex

%/========== Acknowledgements ==========/%

\begin{acknowledgements}
    Manfredi, you are \textit{without a doubt} the greatest supervisor, the
    best mentor, and one of the most wonderful people in the world.
    Knowing you, you'll likely respond to that by saying 
    ``thank you Adan, you're too kind", and then proceed to laugh uncomfortably before
    changing the topic.
    But it's true, and I don't know how else to tell you how lucky and grateful I am to
    have you in my life.

    These two years of doing research together were the most fun I've ever
    had; yet they've also been fraught with death, sorrow, and the most
    devastating heartbreak I've ever experienced.
    Through it all you've been nothing but kind, honest, generous,
    compassionate, understanding, and so incredibly humble.
    You work unbelievably hard in the background to ensure your students succeed, and you
    ask for so little in return.
    My name is on the title page of this thesis, but if not for the required
    academic formality I would change the authorship to read
    ``Adan Moran-MacDonald, who couldn't have done any of this without the love
    and support of his supervisor and dear friend, Manfredi Maggiore."

    Manfredi: you are my hero.

    % TODO: do the rest of the acknowledgements: parents, SCG friends, other
    % friends, committee

\end{acknowledgements}

%/========== /Acknowledgements ==========/%
% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :