### 2021-09-17 latex/nserc/contributions.tex
First draft of the contributions document

%! TEX program = lualatex

\documentclass[a4paper,12pt]{article}

%/========== Preamble ==========/%
%----------- Set Main Font ----------%
\usepackage{times} % Times New Roman

%----------- Choose Margins ----------%
\usepackage[a4paper,margin=1.87cm]{geometry}

%---------- Bibliography Style ----------%
\usepackage[style=ieee,backend=biber]{biblatex}
\addbibresource{bib.bib}

%----------- Math Packages -----------%
\usepackage{mathtools}

%---------- Remove indenting from paragraphs ----------%
% Also makes each new paragraph start with an vertical space.
\usepackage[parfill]{parskip}

%---------- Add name to Header ----------%
\usepackage{atbegshi,picture}
\AtBeginShipout{\AtBeginShipoutUpperLeft{%
        \put(\dimexpr\paperwidth-5.5cm\relax,-1.2cm){\makebox{\setlength{\fboxrule}{0pt}\framebox{Adan
        Moran-MacDonald}}}%
}}

%/========== Main Document ==========/%
\begin{document}

%---------- Title -----------%
\subsection*{Part I: Contributions to Research and Development}
\subsubsection*{d. Non-peer-reviewed Contributions}
\textbf{Moran-MacDonald, A.} \& Alajaji, F. (2018). The R\'{e}nyi Deterministic
Information Bottleneck and Geometric Clustering. NSERC summer research project,
Queen's University.

\subsection*{Part II: Most Significant Contributions to Research \& Development}
\textit{Describe your role in research, clarify your contribution, provide
    details on significance of tech report, indicate collaborations, discuss
relevance to engineering practice (if appropriate)}
``The R\'{e}nyi Deterministic Information Bottleneck and Geometric
Clustering" is a technical report about a new information-theoretic tool for
clustering datapoints.
My contribution was substantial: I produced the theoretical and simulation
results (with lots of advice from my supervisor, Fady Alajaji), and then I wrote
the report in its entirety.
%In the report, I summarized the existing theory on the
%\textit{information bottleneck} (IB) and its variant, the \textit{deterministic
%information bottleneck} (DIB), which are tools from information theory that
%have been touted as a rigorous foundation for Deep Neural Networks. 
%The DIB in particular is useful for clustering data.
%Its equations are defined using an information-theoretic concept called
%``Shannon's Mutual Information" (SMI).
%In this report, Alajaji suggested we extend the equations by using a more
%general concept called ``R\'{e}nyi's Mutual Information" to get the aptly-named
%\textit{R\'{e}nyi's deterministic information bottleneck} (RDIB).
%I compared the efficacy of the DIB and RDIB for clustering data, and we found
%that the RDIB is more robust when the probability distribution of the data model 
%does not match the true distirbution which generated the data.
This report was used by two undergraduate groups at Queen's University for their
honours theses in 2019. 
It also motivated further research, which was published in [1].

[1] J.J. Weng, F. Alajaji and T. Linder, \textit{An information bottleneck
problem with R\'{e}nyi's Entropy}, Proceedings of the IEEE International
Symposium on Information Theory, Melbourne, Australia, July 2021.

\subsection*{Part III: Applicant's Statement}
\subsubsection*{Research Experience}
\textit{Describe scientific and engineering abilities you gained in past
    research experience, including projects, honours thesis, coop reports. If
    you have work experience, discuss relevance of that to your proposed field
of study and what benefits you gained.}
The following list outlines the scientific and engineering abilities I gained in
my research and professional roles.
\begin{itemize}
    \item \textbf{MASc Research, University of Toronto, 2018 -- 2021}
        In my master's degree, I learned several important research skills.
        My supervisor taught me to formulate research questions
        correctly; I learned the tenacity and dedication required to solve a
        complex problem; I found creative solutions by attacked problems from
        multiple directions; I learned to read papers, perform a literature
        review, think critically about the literature, and to apply existing
        results to my research; I improved my time management skills, practiced
        explaining my work (both orally and in writing), and improved my
        presentation skills for the thesis defense. Finally, I learned to write
        research papers and a thesis, which is important for communicating my
        work to the engineering community.
        I also learned several engineering skills, including: advanced control
        techniques; simulating physical systems in MATLAB; and implementing
        control systems in hardware.

    \item \textbf{Radar Developer, University of Toronto, 2018}
        As part of a course project, I learned useful engineering skills by
        writing software for a radar device.
        First, I improved my project management skills by keeping our group
        meetings organized, and by separating the tasks in such a way
        that everyone was learning something from the project.
        I learned to read hardware documentation; to program a hardware
        device by implementing the documentation in software;
        and to validate the claims of the hardware capabilities for myself,
        i.e., to never blindly trust documentation.

    \item \textbf{Undergraduate Thesis, Queen's University, 2017 -- 2018}
        My undergraduate thesis was my first time doing a research project.
        I learned to do collaborative application-based research; that is, I
        worked with my peers to find an application for existing theoretical
        results. During this project I also learned to perform a basic
        literature review, and to manage deadlines with a supervisor.
        In terms of engineering skills, I learned to write reports in LaTeX,
        and my thesis group worked together on code optimization. We learned to
        convert equations into matrix form to make use of hardware speed; we
        parallelized our code to take advantage of multithreading; we ported
        MATLAB code to C++ to achieve faster algorithms; and we created a GUI so
        we could test our code quickly and easily.

    \item \textbf{Software Developer, Neptec Technologies Crop, 2016 -- 2017}
        My experience at Neptec was incredibly beneficial.
        I worked on several embedded software projects, including a control
        systems project. The exposure to various areas of electrical and
        computer engineering is what convinced me to pursue research in the
        first place.
        I learned to work on large software projects, where multiple people work
        together on the same topic.
        I also became extremely proficient at software development, a skill
        which has served me well in my research experiments on physical robots.

\end{itemize}

\subsubsection*{Relevant Activities}
\textit{Describe professional, academic, extracurricular activities that
    demonstrate communication, interpersonal and leadership skills, e.g.
    teaching, mentoring, managing projects, volunteer work, committee work,
participation in organizations and clubs, or work experience}.

Over the course of my career, I have enjoyed several extracurricular roles which
developed my communication, interpersonal, and leadership skills.
What follows is a brief description of some of these roles, along with a summary
of what I learned.
\begin{itemize}
\item \textbf{Lionel Massey Fund Co-chair, Massey College, September 2021 -- Present}.
    I am one of six co-chairs for the Lionel Massey Fund (LMF) -- the social
    committee for Massey College, a graduate residence at the University of
    Toronto.
    In managing 13 orientation events for 150 members of the college, I have
    learned to explicitly communicate my needs with others, and to ask pointed
    questions to understand their needs.

\item \textbf{Teaching Assistant, 2017 -- Present}.
    I have been a teaching assistant for 14 engineering courses in the past 5 years.
    Each course I teach improves my communication skills, and I've learned
    many ways of making course material fun and engaging.
    I focus most on student participation by finding making them feel
    comfortable asking questions.

\item \textbf{Jiu-Jitsu Instructor, 2019 -- Present}. 
    I am an instructor at the University of Toronto Jiu-Jitsu club. I teach
    people martial arts while fostering a safe, friendly atmosphere that
    encourages learning through action.
    This role has refined my communication and interpersonal skills:
    it has taught me to create learning plans for my students;
    break down difficult concepts into manageable chunks; 
    and to reward mistakes, because the best way to learn is to try, fail, and
    always try again.

\item \textbf{Science Payload Designer, Queen’s University Satellite Design
    Team, 2017 -- 2018}. 
    In this role, I taught C++ and information theory to the team’s software
    developers. I learned project management skills by organizing
    the satellite codebase, and I took on a supportive role for my peers rather
    than writing the software myself.

\item \textbf{Head Programming Mentor, FIRST Team 2809, 2013 -- 2015}. 
    I mentored high school students in Object Oriented programming for robotics
    software. Together we programmed a robot in under 6 weeks and placed 3rd at
    the Ontario provincial competition in 2014.
    In this role I learned to lead my students by giving them agency over the
    project, and to communicate with people from all over the world in
    high-stress situations.

\end{itemize}

I take pride in my teaching, communication, and leadership abilities.
Obtaining an NSERC scholarship will provide me with the financial liberty to
participate in more volunteer roles, where I can better myself while giving back to my
community.

\end{document}

% vim: set tw=80 ts=4 sw=4 sts=0 et ffs=unix :