### 2023-07-07 3.FINAL_VERSION/BIOS/maggiore.tex
Final, accepted version sent to the journal

Manfredi Maggiore received the "Laurea" degree in Electronic
Engineering in 1996 from the University of Genoa and the PhD degree in
Electrical Engineering from the Ohio State University, USA, in
2000. Since 2000 he has been with the Edward S. Rogers Sr. Department
of Electrical and Computer Engineering, University of Toronto, Canada,
where he is currently Professor. He has been a visiting Professor at
the University of Bologna (2007-2008), and the Laboratoire des Signaux
et Syst\`emes, Ecole CentraleSup\'elec (2015-2016). His research
focuses on mathematical nonlinear control theory, and relies on
methods from dynamical systems theory and differential geometry.