### 2023-07-07 3.FINAL_VERSION/BIOS/moran-macdonald.tex

Adan Moran-MacDonald graduated from Queen's University, Canada, in 2018,
with a BASc. in Mathematics Engineering and a subspecialty in Systems and Robotics.
He then attained an MASc. in Electrical and Computer Engineering under the 
tutelage of Manfredi Maggiore at the University of Toronto, Canada, in 2021, 
with a focus on Systems Control.
Adan's academic research interests lie in the field of geometric nonlinear control
theory and its applications to biologically-inspired robotics.
He is now an Automation Innovator at Artificial Futures in Toronto, Canada,
where he specializes in research and development for human-centric AI, robotics,
and automation applications.